BareTag Tool-Tracker\\ {\footnotesize {Team 3 - University of Massachusetts Amherst Senior Design Project 2025 }}

### 2024-10-21 draft.tex
proof read ken's writing

@@

\section{Introduction}
On constructions sites tools are constantly being misplaced causing workers to spend extra time searching and waiting before 
returning to their assignments. Furthermore, in recent years research has 
shown that theft on constructions sites has built up costs for construction 
companies and slowing down projects, very detrimental for small companies. 
Many different approaches have been taken to solve this 
growing issue.

@@

Another comparable solution we evaluated is a commercial system called 
GoCodes, which specifically addresses the issue of tool tracking. This 
system employs continuous GPS tracking to monitor valuable equipment 
within a geofence, helping businesses prevent the loss or theft of tools. 
While it offers effective on-site tracking, off-site tracking relies on 
individuals scanning QR codes placed on the tools. If a user scans the 
QR code, the item's location outside of the geofence can be updated [5]. 
However, this approach has a significant flaw — no tool thief would 
voluntarily scan the QR code, and they could easily remove the codes 
from the tools, eliminating any trace of the theft. Although GoCodes 
provides a workable solution, it has several drawbacks that our system 
aims to overcome. For one, GPS tracking is power-intensive, often requiring
frequent battery replacements for the trackers. Additionally, GPS 
tracking needs an unobstructed line-of-sight to function efficiently, 
which can be compromised in dynamic construction environments. 
Furthermore, our system is designed to operate in very rural construction 
areas where satellite connectivity may be limited. Unlike GPS, which 
requires a strong satellite connection for accurate tracking, 
ultra-wideband (UWB) technology, which our system uses, does not rely on 
satellites, making it more reliable in such environments.

A common solution and a more reliable approach is using the Apple AirTags.
In the Apple AirTags, there are BLE modules that help it communicate 
with other iPhones to create the Find My network. Additionally, it has 
UWB modules that allow the user to get a more accurate location of where 
the tag is once the AirTag is close enough to the user’s iPhone [6]. This 
creates a well-rounded system to let users get a general idea of where 
their items are with a battery life of up to a year. The difference 
between the BareTag and the AirTag is that the BareTag is used for 
accurately tracking high-value items within the set UWB perimeter. 
BareTag emphasizes the use of the UWB module so construction workers 
can easily find their tools, along with an additional BLE module for 
the Find My Network feature outside of the set perimeter. Unlike Airtag, 
our system aims to not only track and retrieve items but also to accurately 
and reliability track tools to optimize efficiency.

The final solution takes a similar approach to BareTag. Pozyx tracks 
items indoors by installing UWB anchors throughout the building, which 
communicate with UWB tags. These anchors are height-adjustable and 
plugged into walls. Once configured and registered in the system, 
items can be geofenced, and the collected data can be used to optimize 
workflow and labor efficiency [7]. While Pozyx offers an effective 
indoor tracking solution, it is less efficient in outdoor environments. 
In locations like construction sites, where power availability may be 
limited, anchor placement becomes a significant challenge. Additionally, 
even though Pozyx’s UWB tags are advertised to have a battery life of over 
five years, they are bulky and lack a low-power option. In contrast, 
BareTag’s anchors come with pre-installed communication systems and 
built-in batteries, making the system more portable and scalable. 
Furthermore, BareTag’s tags feature a built-in accelerometer with an 
automatic sleep function to maximize power efficiency. This not only 
extends battery life but also reduces the frequency of battery 
replacements, saving time and resources. In summary, BareTag's anchors 
do not rely on external power sources, and the tags are equipped with a 
low-power solution, making them more suitable for outdoor and construction 
site use compared to Pozyx’s solution.

@@

\section*{Appendix}
\setcounter{subsection}{0}
\subsection{Design Alternatives}
Before settling on our current design, we considered many alternatives. 
In order to make our user-interface have real world terrain displayed 
while tracking our tags, we need to find the coordinate positions of our 
anchors. In order to do this, we originally considered putting GPS modules 
on our anchors, but decided instead on creating a mobile app that can be 
used to find the GPS of your phone when setting up the anchors. We decided 
on this in order to save money, as individual GPS modules range from 
\$20-\$40. We also considered purchasing a separate MCU, LoRa, and UWB 
module to design our anchor PCBs. We found this design too complex and 
instead decided on the same module as we are using in the Tags. We decided 
on this current design because we want the tags and anchors to have the 
same technology to be able to program them with the same libraries in 
order expedite software development. Batteries have been a prominent 
part of our design that we have put a lot of effort into deciding on. 
Because we want the Tags to be small enough to go inside of tools without 
manufacturers having to add room for them, we decided on very small 
batteries inside the tag. We also considered doing energy harvesting 
for the anchors to increase the battery life of them, but have decided 
on large batteries for the anchors because they are less size constrained 
compared to the Tags. In our original plan, we were going to backhaul our 
data to a cloud database in order for processing. After some consideration, 
we decided on sending the data to a self-hosted database to have increased 
control over the hardware, software, and configurations, which allows us to 
customize our database to meet our project’s needs. Having our own 
self-hosted database can also allow us to implement our own security 
measures, will be less expensive, and will have greater performance 
optimized for our specific workload.

>>>>

\and 
\IEEEauthorblockN{Jeremy Gummeson}
\textit{Professor, ECE}\\
\textit{Advisor}

@@

\section{Introduction}
Modern construction sites heavily rely on the usage of battery operated 
power tools, allowing construction workers to dynamically move throughout 
a site without the need to route cables and power lines for their 
equipment. This allows individual workers, as well as entire 
construction firms, to be more efficient, eliminating the need to set 
up power infrastructure at every new site, or re-work power infrastructure 
when a site changes. However, this flexibility and ease of use create new 
challenges to construction site efficiency. First of all, battery-powered 
tools are often smaller and easier to misplace compared to their corded 
predecessors, greatly reducing a firm’s overall efficiency. Second, 
battery-powered tools are very valuable, with many cordless power tools 
costing \$800+, greatly increasing the incentive for thieves to steal 
cordless tools and accessories. In order to prevent inefficiencies caused 
by misplaced tools and to deter common tool theft, a precise and scalable 
tracking system needs to be implemented for the construction industry. 
With this project, we propose a system that combines functionality of 
existing item tracking and inventory management solutions, into a highly 
precise and scalable solution specifically tailored to the construction 
industry. While this solution is focused on preventing tool misplacement 
and theft, we have also designed the system to be flexible, allowing 
users to utilize the system to track and monitor equipment, materials, 
or even individuals.

@@

Another comparable solution we evaluated is a commercial system called 
GoCodes, which specifically addresses the issue of tool tracking. This 
system employs continuous GPS tracking to monitor valuable equipment 
within a geofence, helping businesses prevent the loss or theft of tools. 
While it offers effective on-site tracking, off-site tracking relies on 
individuals scanning QR codes placed on the tools. If a user scans the 
QR code, the item's location outside of the geofence can be updated [5]. 
However, this approach has a significant flaw — no tool thief would 
voluntarily scan the QR code, and they could easily remove the codes 
from the tools, eliminating any trace of the theft. Although GoCodes 
provides a workable solution, it has several drawbacks that our system 
aims to overcome. For one, GPS tracking is power-intensive, often requiring
frequent battery replacements for the trackers. Additionally, GPS 
tracking needs an unobstructed line-of-sight to function efficiently, 
which can be compromised in dynamic construction environments.

Another popular solution is the Apple AirTag.
The Apple AirTag, uses Bluetooth Low Energy (BLE) advertisements to connect
with devices in the Find My network (iPhone, iPads, and Macbooks). 
Additionally, it uses UWB to allow users to get a more accurate location of where 
the tag is once the AirTag is close enough to the owner's iPhone [6]. This 
creates a well-rounded system to let users get a general idea of where 
their items are with a battery life of up to one year. The difference 
between the BareTag and the AirTag is that the BareTag is used for 
accurately tracking high-value items within the set UWB perimeter. 
BareTag emphasizes the use of the UWB module so construction workers 
can easily find their tools, along with an additional BLE module for 
the Find My network features outside of the set perimeter. Unlike Airtag, 
our system aims to not only track and retrieve items but also to accurately 
and reliability track tools to optimize efficiency.

The final solution, Pozyx, is most similar to the BareTag. Pozyx tracks 
items indoors by installing UWB Anchors throughout the building, which 
communicate with UWB Tags. These anchors are height-adjustable and 
plugged into walls. Once configured and registered in the system, 
items can be geofenced, and the collected data can be used to optimize 
workflow and labor efficiency [7]. While Pozyx offers an effective 
indoor tracking solution, it is less efficient in outdoor environments. 
In locations like construction sites, where power availability may be 
limited, anchor placement becomes a significant challenge. In contrast, 
BareTag’s Anchors are battery powered and easy to reconfigure, 
making the system more portable and scalable. 
Furthermore, BareTag’s Tags feature a built-in accelerometer with an 
automatic sleep function to maximize power efficiency. This not only 
extends battery life but also reduces the frequency of battery 
replacements, saving time and resources. In summary, BareTag's Anchors 
do not rely on external power sources, and the Tags are equipped with a 
low-power solution, making them more suitable for outdoor construction 
site use compared to Pozyx.

@@

\section*{Appendix}
\setcounter{subsection}{0}
\subsection{Design Alternatives}
Before settling on our current design, we considered many alternatives. 
In order to make our user-interface have real world terrain displayed 
while tracking our tags, we need to find the coordinate positions of our 
anchors. In order to do this, we originally considered putting GPS 
modules on our anchors, but decided instead on creating a mobile app 
that can be used to find the GPS of your phone when setting up the 
anchors. We decided on this in order to save money, as individual GPS 
modules range from $20-$40. We also considered purchasing a separate 
MCU, LoRa, and UWB module to design our anchor PCBs. We found this 
design too complex and instead decided on the same module as we are 
using in the Tags. We decided on this current design because we want 
the tags and anchors to have the same technology to be able to program 
them with the same libraries in order expedite software development. 
Batteries have been a prominent part of our design that we have put a 
lot of effort into deciding on. Because we want the Tags to be small 
enough to go inside of tools without manufacturers having to add room 
for them, we decided on very small batteries inside the tag. We also 
considered doing energy harvesting for the anchors to increase the 
battery life of them, but have decided on large batteries for the 
anchors because they are less size constrained compared to the Tags. 
In our original plan, we were going to backhaul our data to a 
self-hosted database in order for processing. After some consideration, 
we decided that sending the data to a cloud database, such as AWS, 
would be more beneficial for our project. Using a cloud server will 
allow for scalability, decrease in cost, higher reliability, and 
greater performance than a database we could host in-house. The cloud 
database will also be easier for us to use because they come with 
pre-existing features that will help us easily integrate our project with.
